\chapter{Annotations}
I annotate chess games as a means of self-improvement and personal learning. The market offers a wealth of excellent chess literature, much of which is dense with variations. Additionally, there are countless chess games available, and uncovering hidden insights within these games provides a valuable experience for me.

My primary focus is on identifying critical moments and key moves using straightforward methods, rather than getting lost in complex variations. I prefer to analyze games with fewer moves, often avoiding endgames, as they tend to require extensive calculations and are not my main interest.

After pinpointing these critical moments, I verify my findings using computer analysis. There are times when I miss key points, but other times I succeed in my assessments. This process is an integral part of my learning journey.

I intentionally refrain from using computers to evaluate the players, as I believe such an approach will not contribute to my growth as a chess player.

Ultimately, I aspire to discover effective strategies for playing good chess through simple methods, drawing inspiration from the styles of great players like Karpov and Petrosian.

% Games with high value annotation
\input{../cards/chess/annotations/tal_keller_1959}
\input{../cards/chess/annotations/ding_gukesh_game12}
\input{../cards/chess/annotations/lasker_steinitz_1894}
\input{../cards/chess/annotations/spassky_polugaevsky_1961}

\input{../cards/chess/annotations/kasparov_karpov_1984_game_3}
\input{../cards/chess/annotations/spassky_petrosian_1969_game_5}
\input{../cards/chess/annotations/carlsen_ding_2020}
\input{../cards/chess/annotations/ding_nepomniachtchi_2023_game_4}
\input{../cards/chess/annotations/ding_nepomniachtchi_2023_game_6}
\input{../cards/chess/annotations/karjakin_carlsen_2013}
\input{../cards/chess/annotations/ding_aronian_2013}
\input{../cards/chess/annotations/carlsen_lichao_2015}
\
