\section{Boris Spassky - Tigran Petrosian, World Championship Match 1969 Game 5}
\epigraph{My forte was the middlegame. I had a good feeling for the 
critical moments of the play. This undoubtedly compensated for my lack of opening preparation
 and, possibly, not altogether perfect play in the endgame. In my games things often did 
 not reach the endgame!}{\textit{Boris Spassky}}

\keywords{Middlegame, Critical Position, Passed Pawn}

\begin{multicols}{2}
\newchessgame[
 id=main,
 event={Petrosian - Spassky World Championship Match},
 white={Boris Spassky},
 black={Tigran Petrosian},
 round={5}]

\mainline{
    1. c4 Nf6 2. Nc3 e6 3. Nf3 d5 4. d4 c5 5. cxd5 Nxd5 6. e4 Nxc3
    7. bxc3 cxd4 8. cxd4}
    
\chessboard [
    markstyle=border,
    linewidth=0.05em,
    markfields={c8},
]

I don't like Black's position. White already has a static advantage
because his pawn formation is better and Black has a passive
light square bishop.

On the other hand, White has some natural moves to develop:
\symbishop c4, O-O, \symrook ad1, \symrook fe1 etc.

It is however difficult to recommend any dynamic play for Black here.

\mainline{8... Bb4+ 9. Bd2 Bxd2+ 10. Qxd2 O-O 11. Bc4
    Nc6 12. O-O b6 13. Rad1 Bb7 14. Rfe1} 
    
\chessboard [
    pgfstyle=straightmove,
    linewidth=0.05em,
    markmove={d4-d5},
]

\mainline {14... Rc8 } 

White is planning the d5 thrust to create a central passed pawn.
Black could also tried
\variation[invar]{14... Na5 15. Bf1 Rc8 16. d5 exd5 17. exd5 Qd6 \xskakcomment{Black can defend the ending}}
% I am very happy to find the sequence above!

% 2025. 11.5: I wrote the following during my own analysis. 
% the conclusion is premature because I miss error on Black's next move
% "Now White has a straight forward move, his light square bishop can
% exchange an important defender, Black's light square bishop 
% while the Black's knight can't help too much in the defense."

\chessboard [
	pgfstyle=straightmove,
    linewidth=0.05em,
 	markmove={d4-d5},
]

Time to push the pawn!

\mainline[level=1]{15. d5}

\chessboard
% 2025. 11.5: In my analysis, I saw the knight stuck at a5 for a long time.
% I failed to doubt this move although the move per se was ok.
% 2025. 11.5: I failed to pay attention to the critical position. The move itself was ok. 

Black could also \variation[invar]{15... Na5}
\variation{16. dxe6 Nxc4? 17. exf7+ Kh8 18. Qxd8 Rcxd8 19. Rxd8 Rxd8 20. e5} White wins.
\variation{16. dxe6 Qxd2 17. exf7+ Kh8 18. Nxd2 Nxc4 19. Nxc4 Rxc4 20. e5 Bc8 21. e6 Bxe6 22. Rxe6} equal.
Petrosian didn't like \variation{16. Bd3 exd5 17. e5! Nc4 18. Qf4} White has an attack. Indeed, Polugaevsky won a brilliant game
against Tal later. Black must find the only defense \variation{18... Rc6}.

\mainline[level=1]{15... exd5 16. Bxd5
    Na5 17. Qf4 Qc7 18. Qf5} 

\chessboard

% I missed the error. So did Kasparov. 
% We must always pay attention to non forced exchanges, since it changed the dynamics

\mainline{18... Bxd5?}

This move is bad because it helps White to push his pawn
without a tempo and create a threat. 
After the text move Black has a desperate position. He cannot defend the d5 passed pawn.
In defense it is always important not
to help the attacker. 

White is threatening to play \symknight g5.
Black must parry this threat first before the second one is created. 
A direct \variation[invar]{18... h6} would have been better and White has only a slight advantage. 

\mainline[level=1]{ 19. exd5 Qc2 20. Qf4 Qxa2 21. d6
    Rcd8 22. d7 Qc4 23. Qf5 h6 24. Rc1 Qa6 25. Rc7 b5}
   
\chessboard
    
\mainline[level=1]{26. Nd4 Qb6
    27. Rc8 Nb7 28. Nc6 Nd6 29. Nxd8 Nxf5 30. Nc6 
}
1-0

\end{multicols}
\subsection*{Lessons Learned}
% Lessons learned:
Always pay attention to critical positions: change of pawn formation  (\variation[invar]{15... exd5}), 
unforced exchange (\variation[invar]{18... Bxd5?}).

Often an unnoticiable automatic move in such positions is wrong. The awareness of
such moments is crucial. Being aware of this, it is then possible to find 
better moves in these positions. Like Spassky said, with a good feeling for the
critical moments of the play, sometimes things do not reach the endgame!

Playing the game with Stockfish, I see the evaluation increases gradually in 
the opening. It is hard to tell which move is wrong.
The intuition is correct, White has an easy opening because he has lots of moves.



% Black should seek dynamic play, but where is it??