\newpage
\section{Drawn Positions: Knight and Pawn}
Knight and pawn win easily against a lone king. 
When the pawn is an h-pawn, there are often exceptions. \index{Drawn Positions!Knight and Pawn}

\begin{multicols}{2}
    \chessboard[
        setfen=N6K/4k2P/8/8/8/8/8/8 b - - 0 1,
    ]

    We know from the pawn endgame the defender can use stalemate to draw the game.
    Here Black must choose to move his king to f7 or f8.

    A knight can only move from a dark square to a light square and vice versa.
    Currently White's knight is on a light square. Obviously White can only move 
    his knight. Black's king moves between f7 and f8. 

    In this position, Black must place his king on the light square f7. 
    When the knight is on a light square, it is still White to move.
    When the knight is on a dark square, it is Black to move.
    The knight can move to a dark square to check, and then Black simply plays \symking f8. 
    White cannot force Black's king to leave the f7 or f8 square to free White's king, so the 
    position is drawn.

    If Black plays ...\symking f8 from the original position, when White moves his 
    knight to a dark square, it is Black to move. For example, if White's knight arrives on d8, 
    then Black cannot play \symking f7 anymore. He must free White's king and loses the game.

    As a rule, \textbf{the king should move to the same color square as the knight.}

    \chessboard[
        setfen=8/6k1/8/8/6n1/8/6Kp/8 w - - 0 1,
        showmover=false,
    ]

    This position is also drawn, no matter who has the move. The White king 
    just moves between g2 and h1. Black cannot move his knight, otherwise his pawn is lost.
    To prevent White from playing \symking g2, Black must move his king to h3, but this 
    leads to stalemate.

    \chessboard[
        setfen=8/8/8/8/5k2/2n4p/7P/6K1 b - - 0 1,
        showmover=false,
    ]

    This position is drawn because it is again impossible for Black to 
    drive the king from the corner or to mate him. The best Black can achieve is stalemate.

\end{multicols}