\chapter*{Introduction}
\addcontentsline{toc}{chapter}{Introduction}
\epigraph{Investing isn't about beating others at their game. It's about controlling yourself at your own game.}{Benjamin Graham}

\section*{Horse Race}

Even a fool knows how to pick a horse: fast runners with less weight to carry.

Most horse race bets are based on pari-mutuel betting. All bettors put their money into a pool, the organizer takes a fee (typically 17\%), and the winner takes the rest.

To make money, you must be right often enough, and right when the odds favor you. Occasionally, few people bet on a favorite horse, creating favorable odds. Then you should bet heavily.

The key insight: what others bet doesn't matter. Pick a good horse when the odds favor you.

\section*{Investment}

Investment resembles horse racing. Focus on picking a good company (a good horse) when the price (odds) favors you. The crowd's enthusiasm or fear doesn't matter—only the company's quality and the price. 

\section*{Good Company}

A good company has an easy-to-understand business and generates good profits. Warren Buffett's investment philosophy centers on this: invest only in businesses you understand. If you can't explain how a company makes money in simple terms, don't invest in it.

A good company has several characteristics. First, it operates a simple, understandable business model. Its products or services are clear, its competitive advantages obvious, and its path to profitability straightforward. Complex businesses with opaque operations more likely hide problems or face unexpected challenges.

Second, a good company has strong profit margins and returns on capital. It generates cash, not just revenue. A company that grows rapidly but burns cash isn't a good company—it's a speculation. Buffett looks for businesses with pricing power—they can raise prices without losing customers. This indicates a durable competitive advantage.

Third, a good company has a sustainable competitive advantage—what Buffett calls a ``moat.'' This could be a strong brand, network effects, economies of scale, or regulatory protection. The moat protects the company from competitors and maintains profitability over time.

Finally, a good company is run by honest, capable management. Managers allocate capital wisely, treat shareholders fairly, and focus on long-term value creation rather than short-term metrics.

You don't need to understand every business. You only need to understand a few good businesses well enough to recognize when they're available at a good price. As Buffett says, ``It's far better to buy a wonderful company at a fair price than a fair company at a wonderful price.''

\section*{Good Price: Mr. Market}

Benjamin Graham introduced the concept of ``Mr. Market'' as a metaphor for the stock market. Imagine you own a small share of a business, and you have a partner named Mr. Market. Every day, Mr. Market offers to either buy your share or sell you more shares at a price he sets.

The catch is that Mr. Market is emotionally unstable. Some days he is euphoric and offers you very high prices, believing the business is worth far more than it actually is. Other days he is depressed and offers you very low prices, convinced the business is worthless. His mood swings wildly based on news, rumors, and his own psychological state—not the actual value of the business.

Mr. Market is there to serve you, not guide you. Don't let his mood swings influence your judgment about the business's true value. When he's euphoric and offers high prices, you can sell. When he's depressed and offers low prices, you can buy. Never let his emotional state dictate your decisions.

Most investors make the mistake of following Mr. Market. When prices rise and everyone is optimistic, they feel compelled to buy, fearing they'll miss out. When prices fall and everyone is pessimistic, they feel compelled to sell, fearing further losses. This is backwards. Buy when prices are low and sell when prices are high—exactly the opposite of what Mr. Market's mood suggests.

The discipline to ignore Mr. Market separates successful investors from the crowd. When everyone sells and prices are depressed, that's often the best time to buy a good company at a good price. When everyone buys and prices are inflated, that's often the best time to sell or avoid buying.

Mr. Market will always be there, offering prices every day. You don't need to transact every day. Transact only when the price favors you—when you can buy a good company below its intrinsic value, or sell when the price exceeds it. The rest of the time, ignore Mr. Market's daily mood swings and focus on understanding the business itself.

\section*{Is Investment Simple?}

Yes, investment is simple in principle: pick a good company at a good price. But it's not easy in practice. The difficulty lies not in understanding the concept, but in executing it consistently.

Can you think independently and resist following the crowd? Can you avoid Deutsche Telekom in Germany during the 1990s bubble, when everyone was buying? Can you avoid dot-com companies during the internet bubble, when prices reached absurd levels? Can you avoid Enron in the early 2000s, when it was celebrated as innovative? Can you avoid Wirecard in the 2020s, when it was Germany's fintech darling?

These companies looked attractive to the crowd at the time, but were terrible investments. The crowd's enthusiasm made them expensive, and their underlying businesses were either flawed or fraudulent. Avoiding these mistakes requires discipline, independent thinking, and the ability to say no when everyone else says yes.

This book is about avoiding these mistakes—recognizing the warning signs that separate good investments from bad ones, and resisting the psychological pressure to follow the crowd.

\section*{Where to Find Information}
Finding information shouldn't be difficult. 
The best official source is quarterly/annual reports and shareholder letters.
From there you can find the financial results (the score), but more interestingly, 
how management thinks. Are they focusing mainly on scores?
What is their core business? What is their strategy for the next years? Are they using 
EBITA to hide real costs?

Homepages are also useful. Done by marketing departments, 
they're often filled with marketing buzzwords. A good company should avoid this mistake.

Forums about working culture are also interesting. People write freely 
about their experiences. Read them and ask yourself: is it a company where you want to work?