\section*{P und R Gruppe}
I heard the story from a colleague over many years, not as a warning, but as a sequence of ordinary events.

When he first signed the contract with P&R Gruppe, his life was changing. He had just become a father, and a few years later, another daughter was born, in 2014 and 2017. He told me the investment was for them. Not for speculation, not for early retirement. Just to set something aside, quietly, while time did the work. He invested several thousand euros, money he did not need immediately, money meant for the future of his children.

P&R fit perfectly into that picture. Shipping containers felt almost symbolic: solid, useful, part of the real economy. The promised return of around seven percent sounded reasonable, not greedy. The contracts were long-term, the language reassuring. He signed, filed the papers away, and went back to diapers, sleepless nights, and work.

Then the payments started. They arrived regularly. Each transfer confirmed that the decision had been right. Life moved on. His daughters grew, one learning to walk, the other learning to speak. The investment required no action, no thought. It simply worked. Over the years, the steady income quietly turned trust into habit.

He never talked about risk. There seemed to be none. The money was “working” while he was busy raising his children. That was the whole point.

In 2018, everything broke at once. P&R filed for insolvency. The word sounded unreal when he first said it to me, as if it belonged to some distant scandal, not to his careful plan. The payments stopped. Letters arrived. Legal terms replaced the comforting regularity of bank transfers.

Only then did he understand that the containers he thought he owned were largely fictional, that the seven percent had been paid with other families’ savings, families just like his, investing for children, for stability, for peace of mind. The money meant for his daughters’ future was suddenly frozen, then mostly gone.

\section*{Process of Elimination}

Chess masters use \vocab{PoE}{Process of Elimination}{Ruling out bad moves, not immediately finding the perfect one. Commonly used in calculation, defense, and endgames} to find the best move.
How it works in practice:
\begin{itemize}
  \item List candidate moves that are natural or forcing.
  \item Eliminate bad moves by concrete reasons.
  \item Compare the remaining moves and choose the one that best fits the position.
\end{itemize}

\section*{The Pursuit of Unhappiness}
\input{../cards/life/watzlawick-pursuit-of-unhappiness}

\section*{Charlie Munger: Avoiding Stupidity}
\input{../cards/life/munger-process-of-elimination}

Both Watzlawick and Munger use the Process of Elimination to make better decisions.
But what are the criteria for bad moves? 
Munger lists practical criteria: businesses
with dishonest management, fragile balance sheets, or economics he cannot understand.

\section*{Invest by Elimination}
Charlie Munger analyzed the history of failures and developed his own list of criteria to avoid.
His list works for him. I need to develop my own because 
I have access to the internet, which he rarely used, and I come from a different background. 
In this book, I list criteria that help identify investments to avoid. 
Different companies exhibit different symptoms, so the list will never be complete. 
As the Germans say, ``the way is the goal.'' Let's get started.


\vspace{2em}
\begin{flushright}
Ming Lu\\
\emph{Aegina, Greece}\\
\emph{January 10, 2026}
\end{flushright}
