\newpage
\section{The King Blockades the Passed Pawn}
The stronger side often has a passed pawn, which 
must be blockaded either by the king or by the bishop.

\vocab{first-defensive-system}{
    The first Defensive System
} {
    The king blockades the enemy passed pawn while 
    the bishop defends its own pawns. 
} blockades the enemy passed pawn with the king while the bishop 
defends its own pawns. This is the basic 
and usually the most secure defensive arrangement.

\begin{multicols}{2}
    \newchessgame[
        setfen=8/1k3p2/1P3Kp1/7p/2b2P2/4B1P1/7P/8 w - - 0 1,
    ]
    \begin{chessdiagram}
        \chessboard
    \end{chessdiagram}

    Black's bishop defends the pawns and the king holds the 
    passed pawn. White cannot make any progress

    \mainline[level=1]{
        1. f5 Bd3
    } \textbf{$=$}.

    Breaking down the first defensive system always involve the creation of a second passed pawn, 
    frequently by a pawn breakthrough. \index{Pawn Breakthrough}

    \newchessgame[
        setfen=8/8/4b1p1/2B13p/5P1P/1pK1Pk2/8/8 b - - 0 1,
        moveid=1b
    ]
    \begin{chessdiagram}[Kotov-Botvinnik 1955]
        \chessboard
    \end{chessdiagram}

    A classic example of using pawn breakthrough to break down the first defensive system.

    \mainline[level=1]{
        1...g5! 2. fxg5 d4! 3. exd4 Kg3 4. Ba3}
    
    \variation[invar]{
        4. Be7 Kxh4 5. g6+ Kg4
    }

    Black's bishop has an excellent position. It protects the b3-pawn 
    and restrains both enemy pawns along the \textbf{one diagonal} a2-g8 (\vocab{one-diagonal-principle}{One Diagonal Principle}{}) \index{Bishop Endgame|One Diagonal Principle}

    White has no counterplay. Black just advances his h-pawn and wins the bishop for it.

    \mainline[level=1]{4... Kxh4 5. Kd3 Kxg5}
    

    \chessboard

    Now Black has a second passed pawn and can win the game.

    \mainline[level=1]{
        6. Ke4 h4 7. Kf3 Bd5+
    }

    White resigned because after \symking f2, Black's king goes after the 
    b3-pawn. The bishop defends the h-pawn while restraining the d-pawn 
    along the c8-h3 diagonal.
\end{multicols}