\chapter{Opposite Colored Bishops} \label{chapter:opposite-colored-bishops}

From chapter \ref{chapter:bishop-and-pawns-endgame} and section \ref{section:bishop-and-pawn-drawn-positions} we already learned some basic ideas:
\begin{itemize}
    \item{
        \vocab{one-diagonal-principle}{One Diagonal Principle}{}
    }

    The bishop fulfills its functions from a single diagonal
    \item{
        \vocab{pawns-in-the-crosshairs}{Pawns in the Crosshairs}{}
    }

    attacking the enemy pawns with the bishop either to force the pawns to advance,
    which aids in the task of their subsequent blockade, or else to tie the king to their defense
    \item{
        \vocab{tail-hook}{Tail Hook}{}
    }

    Tying the king to the rearmost pawn from behind
\end{itemize}

\paragraph{Fortress}
Here it is frequently possible to save oneself even two or three pawns down.

Often in opposite-colored bishop endgames, the fortress is constructed rather than played.
First, it is necessary to determine the configuration of pawns and pieces which 
will render the position impenetrable; only then can we proceed with the calculation of variations which will 
prove whether or not we can attain the desired configuration and whether 
it is impenetrable in fact. 

\paragraph{Pawn Placement}
In opposite-colored bishop endings, the stronger side 
should place his pawns on squares of the opposite color to his bishop, while
the weaker side must, contrary to the general rule, keep his pawns on the same color squares 
as his own bishop. ``A bad bishop protects the good pawns.''

\paragraph{Positional Nuance}
When playing an opposite-colored bishop ending, the 
number of pawns is frequently less important than a small alteration 
in the placement of pieces or pawns. Therefore, we often encounter 
pawn sacrifices.

\begin{multicols}{2}
    \newchessgame[
        setfen=1k6/8/1b2pp2/6p1/2P4p/5K1P/2B5/8 w - - 0 1,
    ]
    \begin{chessdiagram}
        \chessboard
    \end{chessdiagram}

    \mainline[level=1]{
        1. c5 Bxc5 2. Bb3 e5 3. Be6 Kc7 4. Ke4
    }

    White achieves a draw by moving his bishop along the h3-c8 diagonal when 
    he is three pawns down. The final position is a \textbf{fortress}.
    White places his pawns on the same color as his own bishop.

    The bishop defends its pawn at h3 and holds the enemy pawns 
    \textbf{on one diagonal}.
\end{multicols}

\input{opposite-colored-bishops/bishop-and-two-connected-pawns}
\input{opposite-colored-bishops/bishop-and-disconnected-pawns}
\input{opposite-colored-bishops/king-blockades-the-passed-pawn}