\chapter{Chess Mindset}

Alexander Kotov's idea of candidate moves brings order into the 
thinking process. Before calculating, the player first evaluates the position 
and identifies its main demands. Only then are a few logical candidate moves 
selected, based on forcing ideas or clear positional improvements.

Each candidate move is calculated separately and thoroughly, without jumping 
between lines. Kotov stressed that mixing variations leads to confusion and 
repeated calculation. After comparing the results, inferior moves are eliminated, 
and the right move is simply the best remaining candidate. In this method, 
good chess is less about brilliance and more about disciplined, structured thinking.

This systematic approach works perfectly for a computer with a rigid algorithm and 
extensive memory. However, trying to emulate a computer misses the point completely.

\section{The Human Challenge}

Chess playing is energy-consuming. There are countless variations to calculate, 
and for human beings, visualizing a given position is already difficult enough. 
Navigating through different variations and remembering the differences 
costs a great deal of mental energy.

\section{A Mindset-Based Approach}

At the end of the day, we are only interested in finding the right moves. 
Quite often, adopting the correct mindset—using the Process of Elimination, 
the Principle of Choices, and Prophylaxis—makes it possible to find the right move 
without consuming excessive energy. These conceptual tools help us focus our 
calculation on what truly matters, allowing us to work with our human limitations 
rather than against them.

\input{../cards/chess/process_of_elimination}
\input{../cards/chess/principle-of-choices}
\input{../cards/chess/prophylaxis}