\newpage
\section{Carlsbad Structure: Black Kingside Offense}

\begin{multicols}{2}
If Black manages to suppress White's minority attack, he can often switch to a 
powerful kingside offensive himself. \index{Queen's Gambit Declined}

\chessgameinfo{Chess Olympiad Final-A}{Milko Bobotsov}{Tigran Vartanovich Petrosian}{}{1968.10.26}{0-1}
\newchessgame
\mainline[level=1]{
1. d4 Nf6 2. c4 e6 3. Nf3 d5 4. cxd5 exd5 5. Nc3 c6 6. Bg5 Be7
7. Qc2 g6 8. e3 Bf5 9. Bd3 Bxd3 10. Qxd3 Nbd7 11. Bh6 Ng4
12. Bf4 O-O 13. O-O Re8 14. h3 Ngf6 15. Ne5}

\chessboard

In this game we see how Black neutralises White's queenside play and takes over
the initiative with a well-placed knight. The knight heads for the dominant
outpost on d6 via b6 and c4. \index{Carlsbad Structure!Black Knight on d6}

This manoeuvre works because White has no light-squared bishop. If he still had
that bishop, he could simply play \symbishop xc4 when Black plays ...\symknight c4.

\mainline[level=1]{15... Nb6! 16. Bg5 Ne4
17. Bxe7 Qxe7 18. Qc2 Nd6! 19. Na4 Nbc4 20. Nxc4 Nxc4 21. Nc5
Nd6!}

\chessboard

Black has now completely suppressed White's minority attack. The queenside is
under control, and he is free to begin a direct kingside assault.
\index{Carlsbad Structure!Black Kingside Attack}
His plan is quite simple:
\begin{itemize}
    \item{Move the knight to e4}
    \item{Move the rooks behind g-pawn}
    \item{Thrust his g- and h-pawn}
\end{itemize}

On the other hand, White has no clear plan. His position is unpleasant and
passive. Black's heavy pieces on the e-file rule out the typical idea of
setting up a center with f3 and e4, so White can do little more than wait.


\mainline[level=1]{22. Rac1 Qg5 23. Qd1 h5 24. Kh1 Re7 25. Nd3 Ne4 26. Nc5
Nd6 27. Nd3 Qf5 28. Ne5 f6 29. Nf3 Rg7 30. Nh2 Re8 31. Kg1 Ne4
32. Qf3 Qe6 33. Rfd1 g5}

\chessboard

Here Black sacrifices a pawn to open lines against the white king. A strong
attack quickly develops and Black wins without difficulty.

\mainline[level=1]{34. Qxh5 f5 35. Re1 g4 36. hxg4 fxg4
37. f3 gxf3 38. Nxf3 Rh7 39. Qe5 Qc8 40. Qf4 Rf8 41. Qe5 Rf5}

\end{multicols}