\newpage
\section{Doubling Rook on G-File from Hedgehog Position}
The Hedgehog is a pawn formation usually adopted by Black that 
can arise from several openings. Black trades the pawn on c5 for White's pawn on d4, then places pawns on a6, b6, d6, and e6. 
These pawns form a row of ``spines'' behind which Black develops pieces. 
Typically, the bishops sit on b7 and e7, knights on d7 and f6, the queen on c7, 
and rooks on c8 and e8 (or c8 and d8).

\chessboard[
    setfen=r4rk1/1bqnbppp/pp1ppn2/8/8/8/8/8 w HAhq - 0 1
]

\begin{multicols}{2}
\chessgameinfo{Exhibition Game}{B. Fischer}{U.Andersson}{}{1970}{1-0}

\newchessgame[
    setfen=r2qrbk1/1pp3pp/2n1bp2/p2np3/8/PP1PPN2/1BQNBPPP/R4RK1 w - - 2 13,
    moveid=13w
]

\chessboard

White reached a color-reversed Hedgehog.
Here Fischer began his original idea: move the king to h1, push g-pawn to g4,
and double rooks on the g-file.

The Black pawn on f6 allows White to play g5 and open the g-file. 

\mainline[level=1]{
    13. Kh1! Qd7 14. Rg1! Rad8 15. Ne4 Qf7 16. g4 g6 17. Rg3! Bg7 18. Rag1!}
    
\chessboard

White plans \symknight h4 and \symknight f5/g5. Black cannot stop the g5 thrust 
with ...g4 because it would weaken the f5 square.

White won the game in 25 moves.


\chessgameinfo{Kislovodsk}{M.Taimanov}{A.Yusupov}{}{1982}{0-1}

\newchessgame[
    setfen=1qr1r1k1/1bbn1ppp/pp1ppn2/8/2P1P3/1NN1BP2/PP4PP/2RR1BQK b - - 0 18,
    moveid=18b
]

\chessboard

Again we have a Hedgehog (color-reversed from the last game), again with a pawn on f3.
It was hard to find a fresh plan for Black. Yusupov knew Fischer's idea and played confidently.

\mainline[level=1]{18... Kh8 19. Rc2 Rg8! 20. Rcd2 g5! 21. Bd4 Rg6! 22. Nc1 Rcg8! 23. Nd3 Qf8}

\chessboard

Black is clearly planning the g4 thrust. 
\mainline[level=1]{24. Re1?}

White has an important resource because
the Black knight cannot jump to h4 as in the previous game due to the pin from the d4 bishop.

\variation[invar]{24. g4} stops Black's pawn thrust. The position stays equal. 


\mainline[level=1]{24...g4 25. fxg4 e5 26. Be3 Nxg4 27. Nd5 Bd8} Black won in the end.

\end{multicols}
