% Alibaba Group Q3 2025 Earnings Call Transcript - Appendix Content
% This file contains the translated content from Alibaba's September 2025 Quarter Earnings Call
% Source: https://data.alibabagroup.com/ecms-files/1508695866/62b5dc8f-ce28-4c6d-b4d3-ce464910c42c/CN%20Sep%202025%20Q%20-%20Alibaba%20Earnings%20Call.pdf
% Translated from Chinese

\section*{Introduction}

Lydia Liu (Bingnan Liu)

Alibaba Group Investor Relations Director

(Translation)

Hello everyone, welcome to Alibaba Group's September 2025 quarter earnings conference call. Attending today's call are: Alibaba Group Chairman Joe Tsai, Chief Executive Officer Yongming Wu, Chief Financial Officer Toby Xu, and Alibaba E-commerce Business Group Chief Executive Officer Fan Jiang.

Please note that this conference call is being webcast live on the Group's website. A recording of today's call will be uploaded to our website later today.

Now I will briefly introduce the disclaimer. Today's discussion may contain forward-looking statements, particularly regarding our business and financial results. Such statements are subject to various risks and uncertainties that may cause actual results to differ materially from the content of the forward-looking statements. For details, please refer to the safe harbor statement in our press release and investor presentation materials released today.

Please note that certain non-GAAP financial measures will be used during the conference call. For more details on the company's financial measures prepared in accordance with GAAP and the reconciliation between GAAP and non-GAAP measures, please refer to our quarterly results press release.

Now, please let Eddie speak.

\section*{Alibaba Group Business Progress}

Yongming Wu

Alibaba Group Chief Executive Officer

(Original remarks)

Dear investors, welcome to this quarter's earnings conference call.

Over the past quarter, Alibaba Group's performance maintained steady growth. Excluding RT-Mart and Intime, the Group's overall revenue increased 15\% year-over-year. Our continued investment in core businesses is showing results: China e-commerce CMR increased 10\% year-over-year, and Cloud Intelligence Group revenue increased 34\% year-over-year.

Next, I would like to specifically share our latest progress in AI+Cloud and the major consumer platform.

Driven by continued strong AI demand and growth in public cloud revenue, Alibaba Cloud achieved 34\% revenue growth while external commercial revenue also accelerated to 29\%. AI-related product revenue achieved triple-digit year-over-year growth for the ninth consecutive quarter.

In the cloud computing market, we see two major trends accelerating. On one hand, as AI applications accelerate deployment, more and more developers and enterprise customers tend to choose cloud computing vendors with full-stack AI technology product portfolios. On the other hand, customers' depth and breadth of AI usage are simultaneously increasing, significantly driving demand for traditional cloud products such as computing, storage, and databases. These together drove the acceleration of Alibaba Cloud's external commercial revenue growth this quarter.

This quarter, we continued to invest in improving full-stack AI capabilities, covering the complete technology system from high-performance AI infrastructure to AI foundation models, to AI development frameworks. Among these, our released flagship model Qwen3-Max ranks in the global first tier in specialized tests for capabilities such as large model calling Coding to solve real-world problems and Agent tool calling.

Full-stack AI technology capabilities have become our key competitive advantage. Alibaba Cloud's market share has improved across multiple segments: in the hybrid cloud market, Alibaba Cloud has become a core participant with growth exceeding 20\%, exceeding market average growth, and market share continues to improve. At the same time, growth in the financial cloud market also leads the market, with share continuing to improve. In China's AI cloud market, overall share also maintains leadership, exceeding the sum of second through fourth place. Recently, NBA, Marriott, China UnionPay, Bosch and others have all reached AI cooperation agreements with Alibaba Cloud.

Last week, we officially launched the Qwen project. We hope to build the smartest personal AI assistant based on the most advanced model. Currently, the Qwen App public beta version has been online for one week, and new downloads have exceeded 10 million.

The launch of the Qwen project means Alibaba is pushing forward in both AI to B and AI to C directions. In the AI to B field, we hope to be a world-leading full-stack AI service provider, serving the growing AI needs of thousands of industries. In the AI to C field, based on performance-leading models and Alibaba's ecosystem advantages, we aim to create AI-native super applications for C-end users.

On one hand, Qwen3-Max's overall model intelligence level and tool calling capabilities have reached global leading levels. Combined with Alibaba's rich user scenarios, we see significantly improved user retention on the beta Qwen App. Therefore, we believe it is the right time to fully enter AI to C.

On the other hand, the synergistic effects of AI and Alibaba's ecosystem represent even greater potential. Alibaba is China's only company with leading models plus rich life and consumption scenarios. In the future, Qwen will gradually integrate with e-commerce, maps, local life and other business ecosystem scenarios, hoping to be the first to create the future AI life entry point. Under the combined effect of AI reshaping and ecosystem synergy...

\section*{Q\&A Session}

\subsection*{Question on AI Investment ROIC}

In the current environment, how do you evaluate the ROIC of the entire industry's AI investments related to training and inference?

\subsubsection*{Yongming Wu:}

(Original remarks)

Indeed, as a full-stack AI service provider, we are in a major AI investment cycle. We have our products, or our infrastructure that can be used in various different aspects. However, in terms of priorities, we actually have several internal priority considerations, and I can share some of them.

I think our most critical priority is, first, to ensure the priority of overall base model training, because for our AI infrastructure to gain more customers, or to gain more high-value scenarios, only by continuously iterating and improving our model capabilities can we have the opportunity to unlock or gain more users, as well as unlock more scenarios. After unlocking more high-value scenarios, the entire token consumption volume and token quality, including customers' willingness to pay for these tokens, will gradually increase. So this is a relatively high priority area for us.

In addition, regarding how to use inference, the inference services on Bailian are also a relatively high priority. Because the Bailian platform we built serves global customers, how to make our AI resources available 24 hours a day with high efficiency, peak shaving and valley filling, so that an AI server can better, 24 hours fully loaded, generate more tokens. Expanding the Bailian resource pool is also a relatively high priority area for us.

On the other hand, whether it's our internal AI inference services in various aspects, or external customers' inference services, demand comes after this. But we also have a priority measure in external customer selection. If an external customer uses Alibaba Cloud services comprehensively—for example, they need storage, big data, and our CPU resources—these customers' priority may be higher. If they only purely rent GPU to meet their simple inference service needs, these customers' demand priority will be relatively lower.

Your second question is also very good. We mainly look at this issue from two aspects.

First is from the demand side. On the demand side, we see two very obvious drivers. One is that we see base models, whether foundation models we see, or video generation models, or more future full-modal models, capabilities in these areas are continuously improving. The entire scaling law trend still hasn't hit a wall. We see many breakthroughs from Gemini 3's progress.

As model capabilities improve, AI models can do more and more things, and can adapt to more and more scenarios. In the process of improving capabilities for adapted scenarios or tasks that can be driven, this is also accompanied by the penetration rate of these tasks in various industries. So driven by these two forces, within three years we see AI demand is still very certain.

Under such high growth demand, we look at the supply side. You analysts may have also seen that starting from the second half of this year, globally, whether from wafer manufacturers, to DRAM manufacturers, to storage manufacturers, to CPU and other AI server links, all are in short supply. This wave of shortages is all driven by AI demand. For future expansion cycles of supply chain manufacturers in all aspects, we know they have considerable bottlenecks. But in this, we think at least it's a two- to three-year expansion cycle. Within these two to three years of expansion cycle, we see demand growth, but the supply side is difficult to rapidly improve. So we think within these three years, AI resources are still in a state of supply falling short of demand. And we actually see that in our industry, including ourselves, including the U.S. hyperscalers we see, not only are their new GPUs basically running at full capacity, but even previous generations, or CPUs from three or five years ago, are running at full capacity. So we think within three years, the so-called AI bubble should not really exist.

\subsection*{Question on Market Share Strategy}

In the previous quarter's conference call, management mentioned that BABA currently wants to gain greater market share in China's major consumer market. We also see that BABA has increased investment in the instant retail field over the past few months, and market share has also increased. May I ask management, besides instant retail, which other sub-industries in the major consumer field does BABA think have potential and may increase investment in the future?

\subsubsection*{Fan Jiang:}

(Original remarks)

Because Alibaba has entered many sub-categories in its layout over the past years. In addition to this year, for example, we have made very large investments in instant retail. We have layouts in similar areas such as Hema, including offline supermarket O2O models, including Fliggy, Amap local life.

We now need to better integrate existing businesses, be able to connect our various businesses, so that we have better synergistic effects between businesses, achieving future market share improvement in the major consumer field.

\section*{Closing}

Lydia Liu:

(Translation)

Thank you everyone for participating in today's conference. We look forward to communicating with you again at the December quarter earnings conference call.

[End of transcript]

\vspace{1cm}

\textit{Source: Alibaba Group September 2025 Quarter Earnings Call Transcript, available at \url{https://data.alibabagroup.com/ecms-files/1508695866/62b5dc8f-ce28-4c6d-b4d3-ce464910c42c/CN%20Sep%202025%20Q%20-%20Alibaba%20Earnings%20Call.pdf}}

\textit{Note: This document contains excerpts from Alibaba's September 2025 Quarter Earnings Call Transcript, translated from Chinese.}
