\section*{Risk Factors from Nokia 2008 Annual Report}

We are a global company and have sales in most 
countries of the world and, consequently, our 
sales and profitability are dependent on general 
economic conditions globally and locally. The 
impact of the current global economic turmoil 
and any further deterioration of global economic 
conditions, as well as the related financial crisis, 
on us, our customers and end-users of our 
products, services and solutions, and suppliers 
and collaborative partners may have a material 
adverse effect on our business, results of opera
tions and financial condition. 

Our sales and profitability depend materially on 
the development of the mobile and fixed com
munications industry as well as the growth and 
profitability of the new market segments that we 
target and our ability to successfully develop or 
acquire and market products, services and solu
tions in those segments. If the mobile and fixed 
communications industry develop in an adverse 
manner, or if the new market segments we target 
and invest in grow less or are less profitable 
than expected, or if new faster growing market 
segments emerge in which we have not invested, 
our business, results of operations and financial 
condition may be materially adversely affected.

Competition in our industry is intense. Our failure 
to maintain or improve our market position or 
respond successfully to changes in the competi
tive landscape may have a material adverse effect 
on our business and results of operations.

We need to have a competitive portfolio of prod
ucts, services and solutions that are preferred by 
our current and potential customers to those of 
our competitors. If we fail to achieve or maintain 
a competitive portfolio, our business, market 
share and results of operations may be materially 
adversely affected.
Our business and results of operations, particu
larly our profitability, may be materially adversely 
affected if we are not able to successfully manage 
costs related to our products, services, solutions 
and operations.
Our sales, costs and results of operations as well 
as the US dollar value of our dividends and mar
ket price of our ADSs are affected by exchange 
rate fluctuations, particularly between the euro, 
which is our reporting currency, and the US dollar, 
the Japanese yen, the Chinese yuan and the UK 
pound sterling, as well as certain other curren
cies.
We depend on a limited number of suppliers for 
the timely delivery of sufficient amounts of fully 
functional components, sub-assemblies, soft
ware and content and for their compliance with 
our supplier requirements, such as our own and 
our customers’ product quality, safety, security 
and other standards. Their failure to deliver or 
meet those requirements could materially ad
versely affect our ability to deliver our products, 
services and solutions successfully and on time.
We must timely and successfully develop or 
otherwise acquire the appropriate technologies 
to use in our business. If we fail to develop or oth
erwise acquire these technologies as required by 
the market, or to successfully commercialize such 
technologies as new advanced products, services 
and solutions that meet customer demand, or fail 
to do so on a timely basis, this may have a mate
rial adverse effect on our business and results of 
operations.
Any actual or even alleged defects or other qual
ity issues in our products, services and solutions 
could materially adversely affect our sales, results 
of operations, reputation and the value of the 
Nokia brand.
Our sales derived from, and assets located in, 
emerging market countries may be materially 
adversely affected by economic, regulatory and 
political developments in those countries or by 
other countries imposing regulations against 
imports to such countries. As sales from these 
countries represent a significant portion of our 
total sales, economic or political turmoil in these 
countries could materially adversely affect our 
sales and results of operations. Our investments 
in emerging market countries may also be sub
ject to other risks and uncertainties.

Changes in various types of regulation and trade 
policies in countries around the world could have 
a material adverse effect on our business and 
results of operations.

We are developing a number of new products, 
services and solutions together with other com
panies. If any of these companies were to fail to 
perform as planned, we may not be able to bring 
our products, services and solutions to market 
successfully or in a timely way and this could 
have a material adverse effect on our sales and 
results of operations.

Our sales and results of operations could be ma
terially adversely affected if we fail to efficiently 
manage our manufacturing and logistics without 
interruption or make timely and appropriate 
adjustments, or fail to ensure that our products, 
services and solutions meet our and our custom
ers’ quality, safety, security and other require
ments and are delivered on time and in sufficient 
volumes.

Our products, services and solutions include in
creasingly complex technologies, some of which 
have been developed by us or licensed to us by 
certain third parties. As a consequence, evaluat
ing the rights related to the technologies we use 
or intend to use is more and more challenging, 
and we expect increasingly to face claims that we 
have infringed third parties’ intellectual property 
rights. The use of these technologies may also 
result in increased licensing costs for us, restric
tions on our ability to use certain technologies 
in our products, services and solution offerings, 
and/or costly and time-consuming litigation, 
which could have a material adverse effect on 
our business, results of operations and financial 
condition.

Our products, services and solutions include 
numerous new Nokia, NAVTEQ and Nokia Siemens 
Networks patented, standardized or proprietary 
technologies on which we depend. Third parties 
may use without a license or unlawfully infringe 
our intellectual property or commence actions 
seeking to establish the invalidity of the intel
lectual property rights of these technologies. 
This may have a material adverse effect on our 
business and results of operations.
Our operations rely on the efficient and uninter
rupted operation of complex and centralized 
information technology systems and networks. 
If a system or network inefficiency, malfunction 
or disruption occurs, this could have a material 
adverse effect on our business and results of 
operations.