\chapter{Things Don't Change}

\section{What Doesn't Change: The Bezos Approach}
\input{../cards/life/bezos-what-doesnt-change.tex}

\section{Endgame Theory: A Permanent Investment}
\input{../cards/life/endgame-what-doesnt-change.tex}

\section{Endgame Transitions}
Having knowledge about certain endgames will save us time and energy in calculation. Chess endgames change
from one to another. For example, a rook endgame may transpose into a queen and rook endgame/middlegame,
a pawn endgame, or a queen vs. rook endgame. It is therefore important not only to remember the tricks
in a certain endgame, but also to be aware of or proactively search for endgames that may arise from the
current position.

In chess, we are not only searching for the best move, but also for the best possible transposed position.
We have knowledge about certain endgames. This knowledge is useful when we can connect these
endgames to our current position. For example, pawn endgame positions in which two rook pawns are facing
each other, with one side having a distant passed pawn, are quite common. When we are defending a rook
endgame where the pawns are described as above, the question becomes: can we force a rook exchange, and
is the resulting pawn endgame a draw? If both answers are yes, we have a clear plan for how to defend.

The same principle applies to life. We face unknown situations constantly—new problems, unfamiliar challenges, unexpected circumstances. But we also have knowledge of known situations: patterns we've seen before, problems we've solved, principles we understand. The art lies in recognizing how to transpose the unknown situation into a known one.

When you encounter a difficult negotiation, you may not have faced this exact scenario before. But you know the principles of negotiation: understanding interests, finding common ground, creating value. The unknown situation becomes manageable when you transpose it into the framework of known principles.

When you face a career transition, the specific circumstances are new. But you understand the patterns of change: assessing your skills, identifying opportunities, building relationships. By transposing the unknown into the known, you create a path forward.

The skill is not in memorizing every possible situation, but in developing the ability to see how unfamiliar problems map onto familiar patterns. Just as a chess player learns to recognize when a complex position can be simplified into a known endgame, we learn to recognize when a complex life situation can be understood through principles we already know.