\documentclass[
	11pt,
	a4paper,
	oneside,
]{article}

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\geometry{a4paper, margin=2.5cm}

% Book information for PDF metadata
\hypersetup{
	pdftitle={My Chess Notebook}, 
	pdfauthor={Ming Lu}, 
	pdfsubject={Chess}, 
	pdfkeywords={chess, ideas, games, annotations}, 
	pdfcreator={LaTeX}, 
}

\definecolor{ocre}{RGB}{243, 102, 25} % Define the color used for highlighting

\usepackage{chessboard}
\usepackage{xskak}
\usepackage{epigraph}
\usepackage{multicol}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{etoolbox} % For \ifstrempty command used in chessgameinfo

% Define section colors
\usepackage{sectsty}
\sectionfont{\color{ocre}\Large\bfseries}
\subsectionfont{\color{ocre}\large\bfseries}

% Define a macro for keywords, the keywords are displayed in italic
\newcommand{\keywords}[1]{Keywords: \textit{#1}}

% Include chess tools
\input{../cards/chess/chess-tool}

%----------------------------------------------------------------------------------------
%	DOCUMENT
%----------------------------------------------------------------------------------------

\begin{document}

\title{My Chess Notebook}
\author{Ming Lu}
\date{\today}
\maketitle

\tableofcontents
\newpage

%----------------------------------------------------------------------------------------
%	IDEAS
%----------------------------------------------------------------------------------------

\part{Ideas}
\input{../cards/chess/ideas/ideas}
\input{../cards/chess/ideas/carlsen-vanVely}

%----------------------------------------------------------------------------------------
%	MY GAMES
%----------------------------------------------------------------------------------------

\part{My Games}
\input{../cards/chess/my_games/Ming-Marius-20250117}

%----------------------------------------------------------------------------------------
%	ANNOTATIONS
%----------------------------------------------------------------------------------------

\part{Annotations}
I annotate chess games as a means of self-improvement and personal learning. The market offers a wealth of excellent chess literature, much of which is dense with variations. Additionally, there are countless chess games available, and uncovering hidden insights within these games provides a valuable experience for me.

My primary focus is on identifying critical moments and key moves using straightforward methods, rather than getting lost in complex variations. I prefer to analyze games with fewer moves, often avoiding endgames, as they tend to require extensive calculations and are not my main interest.

After pinpointing these critical moments, I verify my findings using computer analysis. There are times when I miss key points, but other times I succeed in my assessments. This process is an integral part of my learning journey.

I intentionally refrain from using computers to evaluate the players, as I believe such an approach will not contribute to my growth as a chess player.

Ultimately, I aspire to discover effective strategies for playing good chess through simple methods, drawing inspiration from the styles of great players like Karpov and Petrosian.

\input{../cards/chess/annotations/kasparov_karpov_1984_game_3}
\input{../cards/chess/annotations/spassky_petrosian_1969_game_5}
\input{../cards/chess/annotations/ding_gukesh_game12}
\input{../cards/chess/annotations/carlsen_ding_2020}
\input{../cards/chess/annotations/ding_nepomniachtchi_2023_game_4}
\input{../cards/chess/annotations/ding_nepomniachtchi_2023_game_6}
\input{../cards/chess/annotations/karjakin_carlsen_2013}
\input{../cards/chess/annotations/lasker_steinitz_1894}
\input{../cards/chess/annotations/ding_aronian_2013}

\end{document}

