\chapter{Dishonest Management}

\section{Alibaba's Delivery War}
In early 2025, China’s on‑demand delivery market shifted from a two‑player fight into an all‑out instant‑commerce battle. Meituan had long led food delivery, commanding roughly 70\%  of the market with hundreds of millions of daily orders. Alibaba’s Ele.me had been a distant second, and JD.com’s presence was small. But as consumers embraced “one‑hour” delivery for food and everyday goods, JD.com formally entered the fray in April with a dedicated local delivery push that rapidly gained traction in dozens of cities, pushing millions of daily orders and forcing the incumbents to respond aggressively. 

Alibaba, determined not to cede ground again, rebranded and tightly integrated its delivery business with Taobao’s broader commerce ecosystem. Ele.me was merged into a new Taobao Instant Commerce/Taobao Shangou offering that promised sub‑hour delivery for food, groceries, electronics and general retail, using Ele.me’s courier network to fulfill a wide range of orders placed directly inside Taobao. This aimed to turn delivery into a core driver of traffic and repeat engagement on its main marketplace. 

For much of 2025, Alibaba, Meituan and JD.com poured huge subsidies into food and instant deliveries — from free and near‑free orders to steep merchant incentives and massive rider bonuses — in an effort to win share from consumers and build loyalty. One weekend, Alibaba alone reported more than 80 million combined on‑demand orders in a single day, part of nationwide discount campaigns that overwhelmed merchants and upended local business models. 

The competition was fierce and expensive. Meituan, even though it remained the largest platform by volume, saw its profits shrink sharply — one report showed net profit down nearly 89\% in a quarter as costs for discounts and marketing climbed steeply. JD.com also continued to lose money even as its delivery volumes climbed. Alibaba’s overall earnings were hit as well, with profit declines reflecting its heavy investment in subsidies and capacity expansion. In total, analysts estimated that the three companies spent tens of billions of yuan on price cuts and promotions to compete. 

The escalation drew official attention. Chinese regulators repeatedly summoned the platforms and urged them to curb “disorderly competition”, warning that extreme subsidy wars were eroding the rights of merchants and riders and disrupting healthy market development. Regulators and the companies later pledged to pursue “rational competition” and limit excessive price‑based battles. 

By late 2025 the war had no clear winner but left deep marks: Meituan held its market leadership in food delivery, Alibaba grew its instant commerce footprint and daily order volumes, and JD.com carved out a visible third force. All three suffered margin compression and reported steeper losses or weaker earnings. Facing mounting financial pain and regulatory pressure, the platforms signaled a shift away from runaway subsidies toward more sustainable competition, even as they continued to invest in fast delivery and related services
\section{Alibaba's Quarterly Report}

In Alibaba's September 2025 quarter earnings call (see appendix \ref{appendix:alibaba_2025_q3_earnings_call}), the company presented a narrative focused entirely on growth metrics and market share gains, with no discussion of the underlying economics.

The CEO highlighted that ``Alibaba Group's performance maintained steady growth,'' with overall revenue increasing 15\% year-over-year (excluding RT-Mart and Intime). China e-commerce CMR increased 10\\% year-over-year, and Cloud Intelligence Group revenue increased 34\\% year-over-year. Alibaba Cloud achieved 34\\% revenue growth while external commercial revenue accelerated to 29\\%. AI-related product revenue achieved triple-digit year-over-year growth for the ninth consecutive quarter.

The report emphasized market share improvements across multiple segments: in the hybrid cloud market, Alibaba Cloud's growth exceeded 20\%, exceeding market average growth, with market share continuing to improve. Growth in the financial cloud market also led the market, with share continuing to improve. In China's AI cloud market, overall share maintained leadership, exceeding the sum of second through fourth place.

The Qwen App public beta version achieved over 10 million new downloads in just one week. The company emphasized its position as ``China's only company with leading models plus rich life and consumption scenarios,'' positioning itself as the future AI life entry point.

\section{The Numbers}
Some numbers in the last section are suspiciously round. ``over 10 million new downloads'', ``15\% year-over-year'', ``China e-commerce CMR (What is that?) increased 10\% year-over-year''.
I am at least not sure if they are exactly true. Real numbers can be like 13.7\%, but not 15\%.


\input{ebita_as_metric/main}

Throughout the earnings call, the focus remained on percentage increases, market share gains, and growth metrics. The company highlighted EBITA improvements across core segments, with Cloud Intelligence Group showing strong EBITA expansion driven by revenue growth. The narrative emphasized EBITA growth rates that outpaced revenue growth, suggesting improving operational efficiency and margin expansion.

The delivery war investments were framed as strategic market share gains rather than cost drivers. Instead, the narrative focused solely on the scoreboard: revenue growth, market share, EBITA improvements, and user acquisition numbers.

This pattern mirrors the Enron approach: emphasize impressive growth percentages and market leadership claims while avoiding discussion of whether that growth creates value for shareholders. When a company focuses exclusively on the scoreboard (revenue growth, market share) without addressing the playing field (profitability, cash generation, sustainable economics), investors should be skeptical.