\section{Endgame Challenge by Nunn}
\begin{multicols}{2}
    \section*{1}
    \newchessgame[setfen=8/2p5/6K1/8/8/5k2/P7/8]
    \chessboard

    \mythoughts{
        \variation{1. a4 c5} leads to a draw, as both pawns promote.
    }

    So I must play \mainline[level=1]{1. Kf5}.

    \mainline[level=1]{1... Ke3}

    \mythoughts{
        \variation{2. a4} is still a draw, for the same reason as on the previous move. I have no better try.
    }

    \mainline[level=1]{2. Ke5 c6}

    \variation{2... c5 3. Kd5} would be an easy win for White.

    \mainline[level=1]{3. a4 Kd3}

    Both pawns now race to promotion:

    \mainline[level=1]{
        4. a5 c5 5. a6 c4 6. a7 c3 7. a8=Q c2 
    }

    \chessboard

    \mythoughts{
       Queen against a c-pawn is often drawn. If Black's king reaches d2,
       the game is dead drawn, so I must not allow that setup.
    }

    \mainline[level=1]{8. Qd5+ Ke3}
    
    \mainline[level=1]{9. Qg2 Kd3 10. Qg5 Kc3 11. Qc1 Kd3 12. Qb2 Kd2 13. Ke4 Kd1 14. Kd3 c1=Q 15. Qe2# }
    
    \myreflection{
        This is far from a simple position. Up to the queen-versus-pawn phase I
        found the right moves by elimination, but once the queen appeared the winning plan was
        no longer obvious---I did not see the correct continuation.

        A useful reference pattern for such endgames is the fortress with Black's king on d2
        and the pawn still on c2. If Black reaches that setup, the game is usually drawn;
        to play for a win, White must prevent that formation from arising.
    }

    \section*{2}
    \newchessgame[setfen=8/8/1KP5/3r4/8/8/8/k7 w - - 0 1]
    \chessboard

    \mythoughts{
        This is a rare motif: a lone pawn outwits a rook. The first
        move is obvious.
    }

    \mainline[level=1]{1. c7 Rd6+}

    \mythoughts{
        I cannot play \variation{2. Kc5 Rd2}: Black would have perpetual
        checks on the a- and b-files. To escape the checks, White would need to bring the king
        to c3 (otherwise \symrook d2+ wins the c2-pawn). Black could then tuck the king on b1
        and follow up with \symrook c2, picking off the pawn.

        White must therefore march the king down the board while Black keeps checking from behind.
    }

    \mainline[level=1]{
2. Kb5 Rd5+ 3. Kb4 Rd4+ 4. Kb3 Rd3+ 5. Kc2 Rd4 
    }

    The remainder is study-like:
    \mainline[level=1]{
        6. c8=R Rd2+ 7. Kxd2 Kb2 8. Rc3 Kb1 9. Rb3+ Ka1 10. Kc2 Ka2 11. Rd3 Ka1 12. Ra3# 
    }

    \myreflection{
        Black is not obliged to follow this line; for instance, Black could simply allow White
        to promote to a queen. In over-the-board play I would be unlikely to convert
        queen versus rook from scratch, so more study is needed.
    }

    \section*{3}
    \newchessgame[setfen=5K2/2p5/8/2k5/8/2P5/4P3/8 w - - 0 1]
    \chessboard

    \mythoughts{
        Black has a c-pawn, so a queen-versus-pawn endgame is likely to arise.
        I do not yet have a clear sense of when such an endgame is winning and when it is not.

        Let me start by eliminating some moves.

        \variation{1. e4 Kd6}.

        \variation{1. Ke7 Kc4 2. Kd7 c5 3. e4 Kxc3 4. e5} Black's king
        can go to either b2 or d2. White has three more moves before the pawn queens,
        and Black can push to c2 right after White promotes. The position is a draw.

        \variation{
            1. Kf7 Kc4? 2. e4 Kxc3 3. e5 c5 4. e6 c4 5. e7 Kd2 6. e8=Q c3 7. Qd8+ Kc1 8. Qg5+ Kb1 
        }
        A queen generally wins against a pawn on the 6th rank, so no further calculation is needed.

        \variation{
            1. Kf7 Kd5 2. Kf6 Kc4 3. e4 Kxc3 4. e5 c5 5. e6 Kd2 6. e7 c4 7. e8=Q c3 8. Qd8+ Kc1 
        }

    } 

    \myreflection{
        In the two \variation{1. Kf7} lines, it is instructive to compare the resulting positions:
        one is winning, the other is not. The only difference is whether White's own king
        blocks the queen from giving the decisive check on g5.
    }

    \chessboard[
        setfen=3Q4/5K2/8/8/8/2p5/8/2k5 w - - 2 8
    ]

    \chessboard[
        setfen=3Q4/8/5K2/8/8/2p5/8/2k5 w - - 2 9
    ]

    This means the only correct move is \mainline[level=1]{1. Kg7}.

    \variation[invar]{
        1... Kc4
    } simply loses to \variation{2. e4}. We reach a similar position
    when White's pawn queens with Black's pawn still on the 5th rank: White can give
    the check on g5 and is therefore winning.

    \mainline[level=1]{
        1... Kd5
    }

    \chessboard

    \mythoughts{
        Pawn moves are meaningless here. \variation[invar]{2. Kf6} transposes to a position
        we saw earlier, where White cannot give the check on g5. That leaves only one move.
    }

    \mainline[level=1]{
        2. Kf7
    }

    \myreflection{
        The following position is a reciprocal zugzwang: since Black is to move,
        Black loses.
    }

    \chessboard[
        setfen=8/2p2K2/8/3k4/8/2P5/4P3/8 b - - 3 2
    ]

    \variation[invar]{
        2... Kc4
    } again loses, because once White queens he can give the familiar check on g5.

    \mainline[level=1]{2... Kd5 3. Ke7 Ke5 4. Kd7 Kd5 5. Kc7 c4 6. Kd7 Ke5 7. Kc6
    }

    White will pick up the pawn and win.

    \myreflection{
        This is certainly a difficult problem. The queen-versus-pawn endgame is still
        not clear to me; I need to study it further.
    }

    \section*{4}
    \newchessgame[
        setfen=3K4/kp6/p7/1P6/8/8/7P/8 w - - 0 1
    ]
    \chessboard

    \mythoughts{
        After \variation{1. bxa6 b5}, both pawns queen and the position is a draw.

        The first move is obvious.
    }

    \mainline[level=1]{
        1. b6+ 
    }

    Black cannot take the pawn: once White promotes, the new queen also controls
    the promotion square of the a-pawn, and the resulting endgame is lost for Black.

    \mainline[level=1]{
        1... Kb8
    }

    The next moves are obvious.
    \mainline[level=1]{
        2. h4 a5 3. h5 a4 4. h6 a3 5. h7 a2 6. h8=Q a1=Q!
    }

    \chessboard

    \mythoughts{
        Taking the queen would be stalemate.
    }

    Here I fail to find the right move. White's queen should stay on the 8th rank,
    threatening mate with a king move. Black cannot prevent this with a check,
    so the only move is

    \mainline[level=1]{
        7. Qg8 Qa2
    }

    \mythoughts{
       \variation[invar]{8. Qf8 Qa3}
    }
    fails, as moving the queen off the 8th rank allows Black to check on d6,
    and taking the queen is again stalemate.

    The white queen must therefore stay on the 8th rank. The only move is

    \mainline[level=1]{
        8. Qe8 Qa4
    }

    \chessboard

    Going back to h8 makes no progress, because Black can simply repeat by moving
    the queen to a1.

    \mainline[level=1]{
        9. Qe5+ Ka8 10. Qh8
    } 

    Black is in zugzwang and therefore loses:

    \mainline[level=1]{
        10... Qf4 11. Qa1 Kb8 12. Qa7#
    } 

    \section*{5}
    \newchessgame[
        setfen=8/5r2/8/PP6/2K1k3/8/8/8 w - - 0 1
    ]
    \chessboard

    \mythoughts{
        Pawns against rook occur frequently in practical play.

        \variation{
            1. Kc5 Rf5+
        } If White's king steps forward, he cannot escape the checks;
        if he retreats, he makes no progress. In practice this cannot be a good move.

        \variation{
            1. a6 Rc7+
        } Here it feels like Black gets the check for free. \variation{1. b6} at least prevents
        the check and should therefore be better.

        Let us calculate further:
        \variation{
            1. a6 Rc7+ 2. Kb4 Kd5 3. b6 Rc4! 4. Kb3? Rc6
        } Black wins a pawn.

        \variation{
            4. Kb5 Rc1 5. a7 Rb1+ 6. Ka6 Ra1+ 7. Kb7 Kc5
        } This is a typical ``pawns in the crosshairs'' defense: White cannot make any further progress.
    }

    We can actually choose a move by logic: the other moves are bad and we can give
    plausible reasons against them. Such a shortcut is not always reliable, but it
    saves time and is often useful.

    \mainline[level=1]{
        1. b6 Rf1
    }

    \mythoughts{
        \variation{2. Kc5 Ra1 3. Kc6 Rxa5 4. b7 Ra6+ 5. Kb5 Ra1}

        White cannot promote because of the check on b1.
        White also cannot get rid of the rook: \variation{
            5. Kc7 Ra7
        } and
        \variation{
            5. Kd7 Rb6
        }.

        \variation{2. Kc3 Rb1} or \variation{2... Rf8}: Black is even winning.
        This move can likewise be rejected by logic: it does not improve anything,
        while Black can improve his position. In practice I do not need to decide
        whether the move loses outright or merely misses a winning chance---let me
        look for other moves.

        The same holds for any other king move. For example, \variation{2. Kb4 Kd4}
        simply makes White's position worse.

        \variation{2. b7} is rejected because \variation{2... Rc1} followed by \variation{3... Rb1} wins
        the pawn.

        That leaves only one move.
    } 

    \myreflection{
        I made a mistake above: ``\variation{2. b7} is rejected because \variation{2... Rc1} followed
        by \variation{3... Rb1} wins the pawn.''
        In fact:
        \variation{
          2. b7 Rc1+ 3. Kb5 Kd5 4. Kb6 Rb1+ 5. Kc7 Rc1+ 6. Kb6 Rb1+ 7. Kc7 Rc1+ 8. Kd7 Rb1 9. a6 Rb6 10. Kd8 Rd6+ 11. Kc7 Rc6+ 12. Kd8 
        }
    }

    \mainline[level=1]{
        2. a6 Rc1+
    }

    \mythoughts{
        \variation{
            3. Kb3
        } obviously loses a pawn.

        \variation{
            3. Kb5 Kd5 4. a6 Rb1+ 5. Ka6
        } We already know this is a draw due to the pawns-in-the-crosshairs idea.

        Only one move is left.
    }

    \mainline[level=1]{
        3. Kb4 Kd4
    }
    
    There is another beautiful line:
    \variation[invar]{3... Rb1+ 4. Kc5 Rc1+ 5. Kd6 Ra1 6. a7 Kd4}

    \chessboard[
        setfen= 8/P7/1P1K4/8/3k4/8/8/r7 w - - 1 7
    ]

    White cannot play b7 yet, because of 7... \symrook a6+ followed by 8... \symrook xa7.
    Black must oppose kings to prevent White from playing b7.
    White can still march his king up to the eighth rank without the b6-pawn being
    attacked by Black's king.

    \variation{7. Ke6 Ke4 8. Kd7 Kd5 9. Kc8 Kc6 10. b7 
    }

    Now we return to the main line.

    \chessboard

    \mythoughts{
        \variation{
            4. Kb5 Kd5 5. Kb4? Kc6
        } Black is winning.
        \variation{
            5. a7
        } is a draw, as we already know.

        Any king move to the a-file merely wastes a tempo and lets Black threaten
        mate with a king move.

        \variation{
            5. Kb3
        } loses a pawn.

        After eliminating all king moves, only two pawn moves remain.
        I am not sure which one is better.
    }

    \myreflection{
        In such situations, comparison is helpful.
        Black will obviously answer with \variation{6... Rb1} and \variation{7... Kc3}.
        The pawn on b7 cannot promote, so it makes no further progress.
        Black's defense is very instructive, however: chase the king with checks or mate threats.

        \variation[invar]{
            4. b7 Rb1+ 5. Ka4 Kc4 6. Ka3 Kc3 7. Ka2 Rb2+ 8. Ka1 Rb4 9. a7 Ra4+ 10. Kb1 Rb4+ 11. Kc1 Rh4 12. Kd1 Kd3 13. Ke1 Ke3 14. Kf1 Kf3 15. Kg1 Rg4+ 16. Kf1 Rh4 17. Kg1 Rg4+ 
       }

    }

    \mainline[level=1]{
        4. a7 Rb1 5. Ka3 Kc3 6. Ka2 Rb2+ 7. Ka1 
    } Black cannot stop the promotion.

    \myreflection{
        I really enjoyed this problem. Through a process of elimination, I can often
        find the right move. The techniques from this study are useful for practical
        endgames. It also leaves me with one more task: study the queen-versus-rook endgame!
    }

    \section*{6}
    \newchessgame[
        setfen=k7/8/1p6/p1p5/2P4K/8/PP6/8 w - - 0 1
    ]
    \chessboard

    \mythoughts{
        \variation{
            1. a4 
        } must be played, otherwise Black plays ...a4. There 
        is no way to drive Black's king out of the a-file.
    }

    \mainline[level=1]{
        1. a4 Kb7
    }
    
    The result of the game will be decided if White can drive Black's king out of the a-file
    using zugzwang. For such positions often it is confusing to calculate the concrete 
    lines. One may easily forget where the pieces are and whose turn it is.

    \vocab{mined-squares}{Mined Squares}{A pair of squares that are dangerous to step on:
    the side who enters one falls into zugzwang, as if the square had exploded beneath him.
    Both players must therefore avoid these squares. If the opponent does step on one,
    the other side must immediately occupy the paired square to exploit the zugzwang.}
    \index{Zugzwang}
    \index{Mined Squares}

    Let us identify some mined squares while the b-pawn is still on b2.
    Taking opposition on d7 and f7 and then outflanking to the 8th rank does not help White,
    because he has a weakness on the queenside. Imagine White's king on f7 and Black's king
    on d7, with Black to move:
    \variation{1... Kd6 2. Ke8 Ke5 3. Kd7 Kd4 4. Kc6 Kxc4 5. Kxb6 Kb4 } and the position is equal.

    Outflanking to the 5th rank does not help either. Imagine Black's king on d7 and
    White's king on f6: \variation{1. Ke5 Ke7 2. Kd5 Kd7}.

    Anti-opposition, on the other hand, does help. When Black's king is on d7, White's king
    should be on f6. Black cannot play ...\symking f7 because of b3: White removes his
    weakness, takes the opposition, and is ready to outflank.

    So d7 and f6 form a pair of mined squares, and d6 and f5 form another.
    Both d6 and d7 are adjacent to c7; both f5 and f6 are adjacent to g6. So when Black's
    king is on c7, White's king should be on g6, ready to exploit any step Black makes onto
    a mined square.

    That gives us two further pairs: c7/g6 and c6/g5. Extending the adjacent-square argument
    once more, we also find b7/h5.

    The mined squares above exist only while the pawn is still on b2.
    Black cannot sit and wait forever, because otherwise White's king will approach.
    White plays b3 at the moment Black is forced to take the opposition (for instance when
    the kings are on d7 and f7, or c6 and g6).

    \chessboard

    We know that g5 is mined and that Black's king can reach the corresponding square c7,
    so White must avoid g5.

    \mainline[level=1]{2. Kh5 Kb8 3. Kh6 Kb7 4. Kh5 Kc6 5. Kg5 Kd6 6. Kf5 Kd7 7. Kf6 Kd6 8. b3 Kd7 9. Kf7 Kd6 10. Ke8 Ke6 11. Kd8 Ke5 12. Kc7 Kd4 13. Kxb6 }

    \myreflection{
        Thinking in terms of mined squares is quite different from calculating concrete lines.
        It is a useful tool for endgames: ideas such as triangulation, opposition, and
        anti-opposition are all related to mined squares.
        I am very pleased that I worked this out on my own.
    }
\end{multicols}     