\chapter{Toxic Managers}

A discussion on a Chinese tech forum asked: ``What exactly is `Alibaba flavor'? Can I experience it without joining Alibaba?'' An Alibaba employee responded with a collection of quotes that exemplify toxic management behavior. These quotes demonstrate the contradictory, impossible-to-satisfy demands that characterize toxic managers:

\begin{enumerate}
    \item ``I don't care how you do it, that's your business. If you can't even do this, what's your value?''
    
    \item ``Don't ask me about everything, use your own brain. How would I know how to do this? You figure it out yourself.''
    
    \item ``Why didn't you report this to me? What did you report? Why didn't you explain the pros and cons clearly? You have to take full responsibility for this.''
    
    \item ``Don't set so many restrictions for things that haven't happened, don't make so many bad assumptions. Let the business run first. If you don't do it, how will you know if the result is bad?''
    
    \item ``Didn't you think about the risks before starting? I said I don't care? Does saying `I don't care' mean you can ignore it? Is this your job or my job? This year's 3.25, you can't escape.''
    
    \item ``Although you got a 3.25, it doesn't mean you have no value. Opportunities outside aren't great either, so stay for another year. I guarantee you a 3.75 this year.''
    
    \item ``You got a 3.25 last year, so you definitely should get a 3.25 this half-year. It's not enough that you didn't make mistakes. You need to make progress and breakthroughs. So-and-so also made mistakes, but they adjusted quickly and basically compensated for previous errors, so I definitely have to give them a 3.75. Don't always look at others, look at yourself.''
    
    \item ``Although you did very well this year and deserve a 3.75, so-and-so needs a 3.75 more than you. I need to ensure he gets promoted to P8, so this time I can only give you a 3.5.''
    
    \item ``You're a newcomer, make a sacrifice this year, a 3.25 it is.''
    
    \item ``She's a newcomer, how can a newcomer get a 3.25? You take it.''
    
    \item ``It's not about what you think, you need to see what everyone else thinks.''
    
    \item ``Can you just blindly follow everyone's opinions? What about your own thoughts?''
    
    \item ``When you encounter difficulties, find me, I'll help you coordinate.''
    
    \item ``Why do you come to me for everything? What are you doing?''
    
    \item ``Just go report this yourself. What's wrong with P10? Is P10 not a human? Just say I told you to go.''
    
    \item ``Why did you bypass me and report to P10? What are you trying to do?''
\end{enumerate}

These quotes reveal a pattern: toxic managers create impossible situations where employees cannot win. Ask for help, and you're told to think for yourself. Think for yourself, and you're told you should have asked. Report problems, and you're blamed for not preventing them. Don't report problems, and you're blamed for hiding them. Make progress, and someone else's promotion takes priority. Follow the team's opinion, and you're told to think independently. Think independently, and you're told to follow the team.

This is the essence of toxic management: contradictory demands that ensure employees can never satisfy their managers, creating a culture of fear, confusion, and constant failure. For investors, companies with toxic management cultures are red flags—they destroy employee morale, drive away talent, and create environments where good work is impossible.